---
title: "Practice Exam"
author: "Karl Rohe"
date: "2/27/2020"
output: md_document
---

# Practice Exam

This practice exam asks you to do several code wrangling tasks that we have done in class so far. 

Clone this repo into Rstudio and fill in the necessary code.  Then, commit and push to github.  Finally, turn in a link to canvas. 

- lm(performance ~ origin + ... + dest)
- ignore "max_wind_gust"


```{r echo = F}
library(tidyverse)
library(nycflights13)
# flights
# weather
library(lubridate)
```


## 1. Make a plot with three facets, one for each airport in the weather data.  The x-axis should be the day of the year (1:365) and the y-axis should be the mean temperature recorded on that day, at that airport.
```{r}
# check practice.R 

# add column that counts number of day of the year (2013)
weather %>% mutate(day_of_year = yday(time_hour))

# one for each airport in the weather data

## view(weather) to check which airports present

# EWR <- weather %>% mutate(day_of_year = yday(time_hour)) %>% 
#   filter(origin == "EWR")
# 
# JFK <- weather %>% mutate(day_of_year = yday(time_hour)) %>% 
#   filter(origin == "JFK")
# 
# LGA <- weather %>% mutate(day_of_year = yday(time_hour)) %>% 
#   filter(origin == "LGA")


# x-axis = day of the year (1:365)
# y-axis = mean temperature

# 1. change time_hour to day of year
# 2. group by origin and day of year -> ex) EWR-1 one group
# 3. ggplot with three lines for origin

weather %>% mutate(day_of_year = yday(time_hour)) %>% group_by(origin, day_of_year) %>% summarize(meanTemp = mean(temp, na.rm = T)) %>% 
  ggplot()+geom_line(aes(x = day_of_year, y = meanTemp, col = origin))



```

# facet_wrap

```{r Q1_answer}

weatherQ1 <- weather %>% mutate(day_of_year = yday(time_hour)) %>% group_by(origin, day_of_year) %>% summarize(meanTemp = mean(temp, na.rm = T))

# str(weatherQ1)

ggplot(data = weatherQ1) + geom_point(mapping = aes(x= day_of_year, y = meanTemp, color = origin)) + facet_wrap(~origin)

```


## 2. Make a non-tidy matrix of that data where each row is an airport and each column is a day of the year.  
-> each element = average temperature
-> some function with '_' tidy
```{r}

# Check organizing.R

# row = airport
# col = day of the year

weatherQ1 %>% pivot_wider(names_from = day_of_year, values_from = meanTemp)


```


## 3. For each (airport, day) contruct a tidy data set of the airport's "performance" as the proportion of flights that departed less than an hour late. 
```{r}
# check handling.R & organizing.R


# data
## flights
# dep_delay = departure delays in minutes
## dep_delay < 1

# perfData <- flights %>% group_by(origin, day) %>% summarise(count = n(), noDelay = length(dep_delay[dep_delay < 1]), performance = noDelay / count)
# 
# # each (airport, day)
# # airport's performance = flights departed late < 1 hour
# 
# perfData %>% select(origin, day, performance) %>% 
#   pivot_wider(names_from = origin, values_from = performance)


## day = 1:365

##############looks weird !! ##########

perfDataDay <- flights %>% mutate(day_of_year = yday(time_hour)) %>% group_by(origin, day_of_year) %>% summarise(count = n(), noDelay = length(dep_delay[dep_delay < 1]), performance = noDelay / count)
  
Q3.data <- perfDataDay %>% select(origin, day_of_year, performance)

performanceData <- perfDataDay %>% select(origin, day_of_year, performance) %>% 
  pivot_wider(names_from = origin, values_from = performance)

performanceData

```

## 4. Construct a tidy data set to that give weather summaries for each (airport, day).  Use the total precipitation, minimum visibility, maximum wind_gust, and average wind_speed.  
```{r}

# data = weather
# total precipitation = sum(precip)
# minimum visibility = min(visib)
# IGNORE : maximum wind_gust = max(wind_gust)
# average wind_speed = mean(wind_speed)

## day = 1:31
# weather %>% group_by(origin, day) %>% summarise(totalPrecip = sum(precip),minVisib = min(visib), avgWind_speed = mean(wind_speed))

## day = 1:365
Q4.data <- weather %>% mutate(day_of_year = yday(time_hour)) %>% group_by(origin, day_of_year) %>% 
  summarise(totalPrecip = sum(precip), minVisib = min(visib), avgWind_speed = mean(wind_speed))

#############tidy data set???################  
Q4.data 


```

## 5. Construct a linear model to predict the performance of each (airport,day) using the weather summaries and a "fixed effect" for each airport.  Display the summaries.  
```{r}
## data = weather
# (airport, day)
# weather summaries
# fixed effect of each airport

# create weather summaries
## left join
## check playgrounds.R
Q5.data <- Q4.data %>% left_join(Q3.data, by = c("origin", "day_of_year"))

Q5.model <- lm(performance ~ totalPrecip + minVisib + avgWind_speed, data = Q5.data)
summary(Q5.model)
################### include day & origin??? ###########
##################fixed effect??#######################

```

## 6. Repeat the above, but only for EWR.  Obviously, exclude the fixed effect for each airport.
-> fixed effect? 
```{r}

Q6.data <- Q5.data %>% filter(origin == "EWR")

Q6.model <- lm(performance ~ totalPrecip + minVisib + avgWind_speed, data = Q6.data)
summary(Q6.model) 

#########fixed effect...######################
```